Coherent IR-Hyper-Raman Four Wave Mixing Vibrational Spectroscopy

### 2024-07-26 paper.tex
cleanup intro

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of nonlinear spectroscopy methods that form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling and correlations between different vibrational, electronic, and vibronic states. \cite{RN281, RN103, Cho2008} 
Four wave mixing CMDS has resolved anharmonicities, ultrafast dynamics, and other inter- and intra-molecular couplings in numerous systems. \cite{Cho2008, Gaynor2017, Ziegler2018, Ogilvie2019, Bonn2021, RN325}
% DDK: quickly get to vibrational spectroscopy and vibrational-electronic coupling
To target vibrational and electronic coupling, common experiments make use of Raman transitions (DOVE, TSF, CARS, SRS, RISRS, and citations).
Here we explore the less-utilized hyper-Raman transitions in CMDS four wave mixing spectroscopy.

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a+\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$.
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through an inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been utilized in several CMDS studies.\cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024,Wang2021}
% Fully coherent, ultrafast probes of vibronic coupling would yield insight into processes that control ultrafast electronic relaxation in molecular and biological systems. \cite{Bredenbeck2015, Arsenault2021}
% Cho has proposed a parametric, doubly resonant infrared/visible pathway to probe vibronic coupling in isotropic systems. \cite{Cho2001}
% In this method, a single infrared pulse is resonant with a vibrational mode, and a two photon absorption event from a near infrared (NIR) or visible pulse resonant with an electronic or vibronic state is used to generate the four wave mixing signal.
% The non-parametric pathways similar to this method are referred to herein as doubly resonant HDFG (DR-HDFG).
% Methods analogous to DR-HDFG, where two photon absorption from the infrared pulse and one photon absorption from the visible pulse is used to generate output, i.e., stimulated Raman processes, have been reported. \cite{RN301, RN120}

% Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
% The most well known example of such a spectroscopy is vibrational sum frequency generation (vSFG), a three wave mixing technique used to probe vibrational structure and dynamics at interfaces. \cite{Shen1987_CPL, RN224}
Hyper-Raman difference frequency generation (HDFG) spectroscopy and hyper-Raman sum frequency generation (HSFG) spectroscopy are four wave mixing CMDS methods based on direct excitation of a vibrational mode through IR absorption and subsequent scattering to or from the vibrational mode through a hyper-Raman transition.
In HDFG/HSFG, an infrared pulse is resonant with a vibrational mode, while two other excitations stimulate the hyper-Raman scattering process.
The infrared excitation frequency selects the vibrational modes to be stimulated through the hyper-Raman transition.  
The other two excitation frequencies control electronic enhancement for the scattering event.

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares it with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require non-centrosymmetry.
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to FWM cascades (CITE).
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just the vibrational mode and the electronic state (``evx'' coupling). 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as 2-color experiments (with $\omega_2=\omega_3$); vSFG or vDFG setups can be trivially reconfigured to perform HSFG or HDFG measurements, respectively, to investigate bulk dynamics and spectroscopy.

@@

% An experimental complexity in executing Raman based four wave mixing experiments is the demand of three separate input pulses, often at different frequencies.
% For example, in the case of DOVE, two of the three inputs must be tunable infrared pulses. \cite{RN345} 
% Only recently have methods been developed for seamless tuning and scanning of ultrafast optical parametric amplifiers across swaths of frequency space. \cite{RN162, McDonnell2024}
% Since HDFG demands only a single resonance to be scanned, the experiment can be performed using two input pulses: one tunable, infrared OPA to scan across vibrational resonances, and a two photon absorption from the signal process of a separate OPA, or output from the oscillator which pumps the infrared OPA. \cite{Wang2021}
% As such, laboratories experienced in sum frequency generation (SFG) spectroscopy can perform HDFG spectroscopy to understand bulk dynamics and spectroscopy using nearly the same setup.

% Cho has proposed a parametric, doubly resonant infrared/visible pathway to probe vibronic coupling in isotropic systems. \cite{Cho2001}
% In this method, a single infrared pulse is resonant with a vibrational mode, and a two photon absorption event from a near infrared (NIR) or visible pulse resonant with an electronic or vibronic state is used to generate the four wave mixing signal.
% The non-parametric pathways similar to this method are referred to herein as doubly resonant HDFG (DR-HDFG).
% Methods analogous to DR-HDFG, where two photon absorption from the infrared pulse and one photon absorption from the visible pulse is used to generate output, i.e., stimulated Raman processes, have been reported. \cite{RN301, RN120}

%  which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
% HDFG was one of the first infrared CMDS methods to successfully discriminate against non-resonant background.\cite{RN351, RN352}

% Hyper-Raman scattering cross sections are commonly several orders of magnitude smaller than Raman scattering cross sections, making hyper-Raman a difficult process to detect.\cite{RN515, Kelley2010}
% Several authors have proposed parametric six wave mixing techniques as the coherent analogue of hyper-Raman scattering. \cite{Berger1978, Cho1997, Cho1998}
% However, since most six wave mixing methods cascade into four wave mixing processes and their homodyned output would be dependent upon the hyper-Raman scattering cross section, \cite{RN243,Cho1997, Cho1998, Cho2000_Cascade} it is preferable to investigate four wave mixing analogues to hyper-Raman scattering.

@@

% While HDFG showed promise as an upconverted infrared spectroscopy method, it was quickly supplemented by a doubly vibrationally resonant method, doubly vibrationally enhanced (DOVE) spectroscopy, to investigate intra- and intermolecular vibrational coupling. \cite{RN345, RN101, Cho2000}
% As a result, investigations of HDFG stagnated; to date, there are roughly a dozen studies which report on HDFG. \cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024}
% Recent reports highlight non-negligible HDFG interference in DOVE experiments as well as the suitability of HDFG pathways to interpret ultrafast relaxation dynamics in isotropic systems. \cite{Bonn2024, McDonnell2024}
% These reports motivate deeper analysis of HDFG spectroscopy.

Inspired by recent work that demonstrated the presence of HDFG in ultrafast DOVE and terahertz-IR-visible (TIRV) experiments, we investigate the parameters which result in HDFG output. \cite{Cho2000, Bonn2024, McDonnell2024}
This paper is organized as follows.
In \autoref{steadystate}, we first identify the selection rules that drives HDFG in the absence of electronic resonance. 
% Specific, non-zero time-orderings of singly resonant HDFG are identified through a simple analysis of the relevant response functions.
The selection rules of electronically-resonant HDFG are then identified through the Herzberg-Teller dipole expansion.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, the site selective properties and useful experimental aspects of HDFG are discussed in \autoref{quant}.
% In particular, a comparison of vSFG and HDFG output is given, showing that HDFG and vSFG are roughly equally intense.
A method to extract hyper-Raman hyperpolarizabilities ($\beta_{ijk}$) from HDFG spectra is identified and discussed.

>>>>

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of nonlinear spectroscopy methods that form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling and correlations between different vibrational, electronic, and vibronic states. \cite{RN281, RN103, Cho2008} 
Four wave mixing CMDS has resolved anharmonicities, ultrafast dynamics, and other inter- and intra-molecular couplings in numerous systems. \cite{Cho2008, Gaynor2017, Ziegler2018, Ogilvie2019, Bonn2021, RN325}
% DDK: quickly get to vibrational spectroscopy and vibrational-electronic coupling
To target vibrational and electronic coupling, methods commonly make use of Raman transitions (DOVE, TSF, CARS, SRS, RISRS, and citations).
Here we explore the less-utilized hyper-Raman transitions in CMDS four wave mixing spectroscopy.

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a+\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$.
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through an inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been utilized in several CMDS studies.\cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024,Wang2021}

Hyper-Raman difference frequency generation (HDFG) spectroscopy and hyper-Raman sum frequency generation (HSFG) spectroscopy are four wave mixing CMDS methods based on direct excitation of a vibrational mode through IR absorption and subsequent scattering to or from the vibrational mode through a hyper-Raman transition.
In HDFG/HSFG, an infrared pulse is resonant with a vibrational mode, while two other excitations stimulate the hyper-Raman scattering process.
The infrared excitation frequency selects the vibrational modes to be stimulated through the hyper-Raman transition.  
The other two excitation frequencies control electronic enhancement for the scattering event.

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares it with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require non-centrosymmetry.
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to FWM cascades (CITE).
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just the vibrational mode and the electronic state (``evx'' coupling). 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as 2-color experiments (with $\omega_2=\omega_3$); vSFG or vDFG setups can be trivially reconfigured to perform HSFG or HDFG measurements, respectively, to investigate ev coupling in bulk systems.

@@

@@

Inspired by recent work that demonstrated the presence of HDFG in ultrafast DOVE and terahertz-IR-visible (TIRV) experiments, we investigate the parameters which result in HDFG output. \cite{Cho2000, Bonn2024, McDonnell2024}
This paper is organized as follows.
In \autoref{steadystate}, we first identify the selection rules that drives HDFG in the absence of electronic resonance. 
The selection rules of electronically-resonant HDFG are then identified through the Herzberg-Teller dipole expansion.
Not surprisingly, we find the selection rules are nearly identical to those of hyper-Raman scattering.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, the site selective properties and useful experimental aspects of HDFG are discussed in \autoref{quant}.
A new method for extracting hyper-Raman hyperpolarizabilities ($\beta_{ijk}$) from HDFG spectra is identified and discussed.
